\section{Objetivos}


\subsection{Objetivo General}

To develop web application for academic calendar management employing machine learning algorithms for schedule prediction.

\subsection{Objetivos especificos}

\begin{itemize}
    \item Conduct research on machine learning models and techniques relevant to schedule prediction.
    \item Gather and define requirements from stakeholders, and design UI mockups until an accepted version. 
    \item Produce a detailed software architecture and infrastructure design, including a requirements document and architecture diagram.
    \item Implement modules for management of students, teaching staff, and courses, based on the requirements gathered.
    \item Design and implement a schedule management system that includes schedule creation, modification, and conflict evaluation, ensuring real-time conflict detection.
    \item Implement a module to generate reports including: teaching staff requirements, and course schedules, including schedule reports.
    \item write and finalize user and developer manuals for the system, providing guidance for system use and maintenance.
\end{itemize}

% Develop a machine learning configuration module that will enable (through visual scripting) change the parameters of the prediction algorithm. In this situation ML works as a black box which and by adding or removing data inputs the prediction will change. Each ponderation represents weight having direct impact on the eligibility criteria.

% system could create subsets of students (for a course) but should take into account social interactions among students. (roces)


\subsection{Objetivos y acciones}
\begin{longtable}{|p{3in}|p{3in}|}
\caption{Objetivos y acciones} \label{tab:objectivesNactions} \\
\hline

Conduct research on machine learning models and techniques relevant to schedule prediction. &
\begin{itemize}
    \item From the context of machine learning, evaluate most promising algoritms and their programming languages; document research.
    \item Document and execute a PoC of chosen algorithms
\end{itemize}
\\\hline

Gather and define requirements from stakeholders, design UI mockups, and produce a detailed software architecture and infrastructure design, including a requirements document and architecture diagram.&
\begin{itemize}
    \item Research Jala University's public materials such as the Student Catalog and Academic Calendar, to prepare questionaries and have interview with actors that interact in some way with the scheduling process, such as: Registrar Office, Student Services and Academic entity.
    \item Based on interviews data, produce a SysML requirements model and i* model, so that more details of the system can be discovered.
    \item Design UI mockups and have client meetings to define details.
\end{itemize}
\\\hline

Produce a detailed software architecture and infrastructure design, including a requirements document and architecture diagram. &
\begin{itemize}
    \item Research and document architecture candidates that fit system's requirements.
    \item Design the system architecture (e.g. C4) along with it's infraestructure (e.g. sequence diagrams) to high level.
    \item Set up system environments (develop, test, staging and production) automated with continous delivery and deployment pipelines.
\end{itemize}
\\\hline

Implement modules for management of students, teaching staff, and courses, based on the requirements gathered. &
\begin{itemize}
    \item Detailed system design (e.g. class diagram) of the modules.
    \item Implementation of modules, keeping software quality with unit tests and component integration tests.
\end{itemize}
\\\hline

Design and implement a schedule management system that includes schedule creation, modification, and conflict evaluation, ensuring real-time conflict detection. &
\begin{itemize}
    \item Detailed system design (e.g. class diagram) of the module.
    \item Implementation of modules, keeping software quality with unit tests, component integration tests, and subsytem integration test.
\end{itemize}
\\\hline

Implement a module to generate reports including: teaching staff requirements, and course schedules, including schedule reports.
\begin{itemize}
    \item Detailed system design (e.g. class diagram) of the module.
    \item Implementation of modules, keeping software quality with unit tests, component integration tests, and subsytem integration test.
\end{itemize}
\\\hline

Write and finalize user and developer manuals for the system, providing guidance for system use and maintenance.&
\begin{itemize}
    \item Elaboration of user guide manual.
    \item Elaboration of developr guide manual.
\end{itemize}
\\\hline
\end{longtable}
